Notes on Validated Model Counting \\ Version of \today

### 2022-06-15 doc/mc-notes.tex
Start discussion of Tseitin variables
Fix typo in proof

\begin{figure}
  A).  CRAT File Contents
  \begin{center}
  \begin{tabular}{lcllll}
    \multicolumn{4}{c}{File line} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{\tt i}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause}\\
    {\tt 2}        & {\tt p}   & {\tt 4 -2 3}  &            & & Declare $p_4 = \obar{x}_2 \pand x_3$ \\
    {\tt 5}        & {\tt s}   & {\tt 5 2 4}   & {\tt 3 0}  & & Declare $s_5 = x_2 \por p_4$ \\
    {\tt 8}        & {\tt p}   & {\tt 6 -1 5}  &            & & Declare $p_6 = \obar{x}_1 \pand s_5$ \\
    {\tt 11}         & {\tt s}   & {\tt 7 1 6}   & {\tt 9 0}  & & Declare $s_7 = x_1 \por p_6$ \\
    {\tt 14} & {\tt a}  & {\tt 7 0} & {\tt 12 13 7 6 2 1 0} & & Assert unit clause $[s_7]$ \\
             & {\tt dc}  & {\tt 1}  & {\tt 4 5 10 11 14 0} & & Delete input clause \\
  \end{tabular}
  \end{center}
B). Proof Clauses
    \begin{center}
  \begin{tabular}{lcllll}
    \multicolumn{4}{c}{Clause} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause}\\
    {\tt 2} &     & {\tt 4  2 -3 0}  &     & & Defining clauses for $p_4$ \\ 
    {\tt 3} &     & {\tt -4 -2 0}    & & & \\ %%Defining clause for $p_4$ \\ 
    {\tt 4} &     & {\tt -4 3 0}    &  & & \\ %%Defining clause for $p_4$ \\ 
            &     & {\tt -2 -4 0}   & {\tt 3 0}  & &  Mutual exclusion proof for $s_5$ \\
    {\tt 5} &     & {\tt -5 2 4 0}  &     & & Defining clauses for $s_5$ \\ 
    {\tt 6} &     & {\tt  5 -2 0}    &  &  \\  %%Defining clause for $s_5$ \\ 
    {\tt 7} &     & {\tt  5 -4 0}    &  & &  \\  %%Defining clause for $s_5$ \\ 
    {\tt 8} &     & {\tt 6  1 -5 0}  &    & & Defining clauses for $p_6$ \\ 
    {\tt 9} &     & {\tt -6 -1 0}    &  & &  \\  %%Defining clause for $p_6$ \\ 
    {\tt 10} &     & {\tt -6 5 0}    &  & &  \\  %%Defining clause for $p_6$ \\ 
            &     & {\tt -1 -6 0}   & {\tt 9 0}  & &  Mutual exclusion proof for $s_7$ \\
    {\tt 11} &     & {\tt -7 1 6 0}  &  & & Defining clauses for $s_7$ \\ 
    {\tt 12} &     & {\tt  7 -1 0}     & &  \\  %%Defining clause for $s_7$ \\ 
    {\tt 13} &     & {\tt  7 -6 0}     & &  \\  %%Defining clause for $s_7$ \\ 
    {\tt 14} &     & {\tt 7 0} & {\tt 12 13 7 6 2 1 0} & & Assert unit clause $[s_7]$ \\
             & {\tt dc}  & {\tt 1}  & {\tt 4 5 10 11 14 0} & & Delete input clause \\
  \end{tabular}
  \end{center}
  \caption{CRAT file \#1 for formula $x_1 \lor x_2 \lor x_3$, and the resulting set of proof clauses}
  \label{fig:c3:crat}
\end{figure}

@@

    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{\tt i}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause $C_1$}\\
    {\tt 2} & {\tt i}   & {\tt -1 2 0}  &            & & Input clause $C_2$ \\
    {\tt 3} & {\tt p}   & {\tt 4 -2 3}  &            & & Declare $p_4 = \obar{x}_2 \pand \obar{x}_3$ \\
    {\tt 6} & {\tt p}   & {\tt 5 -1 4}  &             & & Declare $p_5 = \obar{x}_1 \pand p_4 = \obar{x}_1 \pand \obar{x}_2 \pand \obar{x}_3$ \\
    {\tt 9} & {\tt a}  & {\tt -5 0} & {\tt 7 8 4 5 1 0} & & Assert $t_1 = \obar{p}_5 = x_1 \lor x_2 \lor x_3$ \\
            & {\tt dc}  & {\tt 1 } & {\tt 3 6 7 0} & & Delete clause $C_1$\\
    {\tt 10} & {\tt p}   & {\tt 6 1 -2} &              & & Declare $p_6 = x_1 \pand \obar{x}_2$ \\    
    {\tt 13} & {\tt a} & {\tt -6 0} & {\tt 11 12 2 0}  & & Assert $t_2 = \obar{p}_6 = \obar{x}_1 \lor x_2$ \\
            & {\tt dc}  & {\tt 2 } & {\tt 10 13 0} & & Delete clause $C_2$\\
    {\tt 14} & {\tt s}   & {\tt 7 5 6}   & {\tt 7 11 0}  & & Declare $s_7 = \obar{t}_1 \por \obar{t}_2$ \\
    {\tt 17} & {\tt a}  & {\tt -7 0}       & {\tt 9 13 0} & & Assert $t_3 = \obar{s}_7 = t_1 \land t_2$ \\
             & {\tt dc}  & {\tt 9}          & {\tt 15 17 0} & & Delete $t_1$\\
             & {\tt dc}  & {\tt 13}         & {\tt 16 17 0} & & Delete $t_2$\\
  \end{tabular}
  \end{center}  
B). Proof Clauses
  \begin{center}
  \begin{tabular}{lcllll}
    \multicolumn{4}{c}{Clause} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause $C_1$}\\
    {\tt 2} &    & {\tt -1 2 0}  &            & & Input clause $C_2$ \\
    {\tt 3} &   & {\tt 4 2 3 0} &     & & Defining clauses for $p_4$ \\
    {\tt 4} &   & {\tt -4 -2 0} &     & & \\
    {\tt 5} &   & {\tt -4 -3 0} &     & & \\

@@

    {\tt 9} &   & {\tt -5 0} & {\tt 7 8 4 5 1 0} & & Assert $t_1 = \obar{p}_5 = x_1 \lor x_2 \lor x_3$ \\
            & {\tt dc}  & {\tt 1 } & {\tt 3 6 7 0} & & Delete clause $C_1$\\

    {\tt 10} &   & {\tt 6 -1 2 0}  &    & & Defining clauses for $p_6$ \\ 
    {\tt 11} &   & {\tt -6 1 0}    &  & & \\
    {\tt 12} &   & {\tt -6 -2 0}    &  & & \\ 
    {\tt 13} &  & {\tt -6 0} & {\tt 11 12 2 0}  & & Assert $t_2 = \obar{p}_6 = \obar{x}_1 \lor x_2$ \\
            & {\tt dc}  & {\tt 2 } & {\tt 10 13 0} & & Delete clause $C_2$\\
             &          & {\tt -5 -6 0} & {\tt 7 11 0}  & & Mutual exclusion proof for $s_7$ \\
    {\tt 14} &   & {\tt -7 5 6 0}  &     & & Defining clauses for $s_7$ \\ 
    {\tt 15} &   & {\tt  7 -5 0}    &  &  \\  
    {\tt 16} &   & {\tt  7 -6 0}    &  & &  \\
    {\tt 17} &   & {\tt -7 0}       & {\tt 9 13 0} & & Assert $t_3 = \obar{s}_7 = t_1 \land t_2$ \\
             & {\tt dc}  & {\tt 9}          & {\tt 15 17 0} & & Delete $t_1$\\
             & {\tt dc}  & {\tt 13}         & {\tt 16 17 0} & & Delete $t_2$\\
  \end{tabular}
  \end{center}  
  \caption{CRAT file \#2A for formula $\phi_2 = (x_1 \lor x_2 \lor x_3) \land (\obar{x}_1 \lor x_2)$, and the resulting set of proof clauses}
  \label{fig:p2-bdd:crat}
\end{figure}

@@

>>>>

\begin{figure}
  A).  CRAT File Contents
  \begin{center}
  \begin{tabular}{lcllll}
    \multicolumn{4}{c}{File line} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{\tt i}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause}\\
    {\tt 2}        & {\tt p}   & {\tt 4 -2 3}  &            & & Declare $p_4 = \obar{x}_2 \pand x_3$ \\
    {\tt 5}        & {\tt s}   & {\tt 5 2 4}   & {\tt 3 0}  & & Declare $s_5 = x_2 \por p_4$ \\
    {\tt 8}        & {\tt p}   & {\tt 6 -1 5}  &            & & Declare $p_6 = \obar{x}_1 \pand s_5$ \\
    {\tt 11}         & {\tt s}   & {\tt 7 1 6}   & {\tt 9 0}  & & Declare $s_7 = x_1 \por p_6$ \\
    {\tt 14} & {\tt a}  & {\tt 7 0} & {\tt 13 12 7 6 2 1 0} & & Assert unit clause $[s_7]$ \\
             & {\tt dc}  & {\tt 1}  & {\tt 4 5 10 11 14 0} & & Delete input clause \\
  \end{tabular}
  \end{center}
B). Proof Clauses
    \begin{center}
  \begin{tabular}{lcllll}
    \multicolumn{4}{c}{Clause} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause}\\
    {\tt 2} &     & {\tt 4  2 -3 0}  &     & & Defining clauses for $p_4$ \\ 
    {\tt 3} &     & {\tt -4 -2 0}    & & & \\ %%Defining clause for $p_4$ \\ 
    {\tt 4} &     & {\tt -4 3 0}    &  & & \\ %%Defining clause for $p_4$ \\ 
            &     & {\tt -2 -4 0}   & {\tt 3 0}  & &  Mutual exclusion proof for $s_5$ \\
    {\tt 5} &     & {\tt -5 2 4 0}  &     & & Defining clauses for $s_5$ \\ 
    {\tt 6} &     & {\tt  5 -2 0}    &  &  \\  %%Defining clause for $s_5$ \\ 
    {\tt 7} &     & {\tt  5 -4 0}    &  & &  \\  %%Defining clause for $s_5$ \\ 
    {\tt 8} &     & {\tt 6  1 -5 0}  &    & & Defining clauses for $p_6$ \\ 
    {\tt 9} &     & {\tt -6 -1 0}    &  & &  \\  %%Defining clause for $p_6$ \\ 
    {\tt 10} &     & {\tt -6 5 0}    &  & &  \\  %%Defining clause for $p_6$ \\ 
            &     & {\tt -1 -6 0}   & {\tt 9 0}  & &  Mutual exclusion proof for $s_7$ \\
    {\tt 11} &     & {\tt -7 1 6 0}  &  & & Defining clauses for $s_7$ \\ 
    {\tt 12} &     & {\tt  7 -1 0}     & &  \\  %%Defining clause for $s_7$ \\ 
    {\tt 13} &     & {\tt  7 -6 0}     & &  \\  %%Defining clause for $s_7$ \\ 
    {\tt 14} &     & {\tt 7 0} & {\tt 13 12 7 6 2 1 0} & & Assert unit clause $[s_7]$ \\
             & {\tt dc}  & {\tt 1}  & {\tt 4 5 10 11 14 0} & & Delete input clause \\
  \end{tabular}
  \end{center}
  \caption{CRAT file \#1 for formula $x_1 \lor x_2 \lor x_3$, and the resulting set of proof clauses}
  \label{fig:c3:crat}
\end{figure}

@@

    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{\tt i}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause $C_1$}\\
    {\tt 2} & {\tt i}   & {\tt -1 2 0}  &            & & Input clause $C_2$ \\
    {\tt 3} & {\tt p}   & {\tt 4 -2 3}  &            & & Declare $p_4 = \obar{x}_2 \pand \obar{x}_3$ \\
    {\tt 6} & {\tt p}   & {\tt 5 -1 4}  &             & & Declare $p_5 = \obar{x}_1 \pand p_4 = \obar{x}_1 \pand \obar{x}_2 \pand \obar{x}_3$ \\
    {\tt 9} & {\tt a}  & {\tt -5 0} & {\tt 7 8 5 4 1 0} & & Assert $t_1 = \obar{p}_5 = x_1 \lor x_2 \lor x_3$ \\
            & {\tt dc}  & {\tt 1 } & {\tt 3 6 9 0} & & Delete clause $C_1$\\
    {\tt 10} & {\tt p}   & {\tt 6 1 -2} &              & & Declare $p_6 = x_1 \pand \obar{x}_2$ \\    
    {\tt 13} & {\tt a} & {\tt -6 0} & {\tt 12 11 2 0}  & & Assert $t_2 = \obar{p}_6 = \obar{x}_1 \lor x_2$ \\
            & {\tt dc}  & {\tt 2 } & {\tt 10 13 0} & & Delete clause $C_2$\\
    {\tt 14} & {\tt s}   & {\tt 7 5 6}   & {\tt 7 11 0}  & & Declare $s_7 = \obar{t}_1 \por \obar{t}_2$ \\
    {\tt 17} & {\tt a}  & {\tt -7 0}       & {\tt 9 13 0} & & Assert $t_3 = \obar{s}_7 = t_1 \land t_2$ \\
             & {\tt dc}  & {\tt 9}          & {\tt 15 17 0} & & Delete $t_1$\\
             & {\tt dc}  & {\tt 13}         & {\tt 16 17 0} & & Delete $t_2$\\
  \end{tabular}
  \end{center}  
B). Proof Clauses
  \begin{center}
  \begin{tabular}{lcllll}
    \multicolumn{4}{c}{Clause} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause $C_1$}\\
    {\tt 2} &    & {\tt -1 2 0}  &            & & Input clause $C_2$ \\
    {\tt 3} &   & {\tt 4 2 3 0} &     & & Defining clauses for $p_4$ \\
    {\tt 4} &   & {\tt -4 -2 0} &     & & \\
    {\tt 5} &   & {\tt -4 -3 0} &     & & \\

@@

    {\tt 9} &   & {\tt -5 0} & {\tt 7 8 5 4 1 0} & & Assert $t_1 = \obar{p}_5 = x_1 \lor x_2 \lor x_3$ \\
            & {\tt dc}  & {\tt 1 } & {\tt 3 6 9 0} & & Delete clause $C_1$\\

    {\tt 10} &   & {\tt 6 -1 2 0}  &    & & Defining clauses for $p_6$ \\ 
    {\tt 11} &   & {\tt -6 1 0}    &  & & \\
    {\tt 12} &   & {\tt -6 -2 0}    &  & & \\ 
    {\tt 13} &  & {\tt -6 0} & {\tt 12 11 2 0}  & & Assert $t_2 = \obar{p}_6 = \obar{x}_1 \lor x_2$ \\
            & {\tt dc}  & {\tt 2 } & {\tt 10 13 0} & & Delete clause $C_2$\\
             &          & {\tt -5 -6 0} & {\tt 7 11 0}  & & Mutual exclusion proof for $s_7$ \\
    {\tt 14} &   & {\tt -7 5 6 0}  &     & & Defining clauses for $s_7$ \\ 
    {\tt 15} &   & {\tt  7 -5 0}    &  &  \\  
    {\tt 16} &   & {\tt  7 -6 0}    &  & &  \\
    {\tt 17} &   & {\tt -7 0}       & {\tt 9 13 0} & & Assert $t_3 = \obar{s}_7 = t_1 \land t_2$ \\
             & {\tt dc}  & {\tt 9}          & {\tt 15 17 0} & & Delete $t_1$\\
             & {\tt dc}  & {\tt 13}         & {\tt 16 17 0} & & Delete $t_2$\\
  \end{tabular}
  \end{center}  
  \caption{CRAT file \#2A for formula $\phi_2 = (x_1 \lor x_2 \lor x_3) \land (\obar{x}_1 \lor x_2)$, and the resulting set of proof clauses}
  \label{fig:p2-bdd:crat}
\end{figure}

@@

%% \subsection{Tseitin Variable Optimization}

%% Many problems cannot be directly encoded in CNF without an exponential
%% blowup.  This can be avoided by introducing auxilliary variables.  A
%% particular approach to these encodings makes use of {\em Tseitin}
%% variables.  The general scheme with these is to let a variable $z$
%% encode a subformula $F$ over input and prior Tseitin variables by
%% adding clauses encoding $z \leftrightarrow F$.  A particular property
%% of this encoding is that if we let ${\cal C}_z$ (respectively ${\cal
%%   C}_{\obar{z}}$) denote the clauses in which literal $z$ (resp.,
%% $\obar{z}$) occcurs
